---
title: "Survey Methodology Group Prj"
format: html
editor: visual
---
#Required Libraries

```{r}
library(haven)
library(tidyverse)

```

Data Cleaning

```{r}
data <- read_dta("ZA7575.dta")

```

```{r}
data_df <- data |> dplyr::select(serialid,isocntry, d70,d71_1,d71_2,d71_3,d72_1,d72_2,polintr, d11,d10,d25,d1,d63,d8, sd2_1,sd2_2,sd2_3,sd2_4,sd2_5,sd2_6,sd2_7,sd2_8,qc5_1,qc5_2, qc5_3,qc5_4, qc7,qc15_1,qc15_2, qc15_3,qc17_3,qc17_4,qc17_5,qc19,qc20, d15a,sd3,qc1_4,qc1_8,qc1_10,d7r3,d62t,opls) 

#changing the name of columns
data_df <- data_df |> rename(Country=isocntry,lifesat=d70,natmat=d71_1, eumat=d71_2,locmat=d71_3,euvoice=d72_1, natvoice=d72_2,age=d11,gender=d10,community=d25,ideo=d1,class=d63,eduyears=d8,ethnic=sd2_1, skin=sd2_2,relig=sd2_3, roma=sd2_4, sex=sd2_5,disab=sd2_6,other=sd2_7, none_min=sd2_8,oncontent=qc5_1,pubdef=qc5_2,campaign=qc5_3,workplace=qc5_4,fightdisc=qc7,suprights=qc15_1, suprelation=qc15_2,supmarriage=qc15_3,ch_marriage=d7r3, schsexor=qc17_3,schtrans=qc17_4,schinter=qc17_5,policy_id=qc19, policy_sex=qc20,occup=d15a,religion=sd3, sexdiscr=qc1_4,transdiscr=qc1_8, interdiscr=qc1_10, inter_usage=d62t,leadership = opls)



```

## Final Assignment

## Libraries

```{r}
rm(list = ls())

```

```{r, results='hide'}

```

## Data

```{r}

```

## DATA CLEANING

Cleaning specific variables to determine which are the more important once for our analysis.


```{r}

```


### Added Variable 

**GDP for Each Country**

Explain why GDP

```{r}

```

**Education Expenditures**

Explain why Education Expenditures

```{r}

```


## EXPLAINING CROSS-COUNTRY DIFFERENCES

In this section, we aim to identify factors influencing support for or opposition to transgender rights, particularly the ability to change civil documents. Our analysis will alternate between individual-level and country-level perspectives 

## Descriptive Analysis

### Gender


```{r}

```

### Political Belief -idk if belief is the right word

```{r}

```


### Religios Belief


```{r}

```



### Contact with Transgender Individual

```{r}

```


### Age


```{r}

```


### Awareness


```{r}

```


### Education



```{r}


```


### The economic situation

```{r}

```

### Support for LGB rights

```{r}

```


### Rural vs. urban


```{r}
d
```


### Country-level economic variables (GDP per capita and unemployment)

```{r}

```


# Conclusion


## Statistical Model



```{r}

```



INDIVIDUAL LEVEL

**Gender** ## EVERYTHING FROM UNTIL OTHERWISE INDICATED CAN BE EDITED TO FIT OUR VARIABLE ##

```{r}


```

**Age**



```{r}

```


**Political_ideology**

*In political matters people talk of "the left" and "the right". How would you place your views on this scale?*

```{r}

```

**Religion_cat**

*Do you consider yourself to be…?*

```{r}


```

**Contact with transgender people (sd1_7)**

*Do you have friends or acquaintances who are transgender?*

```{r}


```

**Age when completing full-time education (d8)**

*How old were you when you stopped full‐time education?*

```{r}


```

**Personal economic situation (d60_ordinal)**

*During the last twelve months, would you say you had difficulties to pay your bills at the end of the month…?*

```{r}

```

**Sexual minority(sd2_5)**

*Where you live, do you consider yourself to be part of a sexual minority?*

```{r}


```

**Place of residence (type) (d25)**

*Would you say you live in a...?*

```{r}

```

**Equal rights for the LGB community (qc15_1_ordinal)**

```{r}


```

**Discrimination against the transgender community (qc1_8_ordinal)**

```{r}


```

COUNTRY LEVEL

**Variables created earlier country level.**


```{r}


```

**adp_general**



```{r}


```

**Country_name** ## END OF VARIABLE EDITS ##



```{r}


```


DATA FINAL COUNTRY

```{r}

```

DATA FINAL INDIVIDUAL


### Regression Analysis Individual Level


```{r}

```


### Multi-level analysis


```{r}

```


### Conclusion



## PREDICTION

Our goal is to develop a model capable of predicting support for transgender individuals changing civil documents. To accurately forecast our target variable, we must first establish a clear definition of the outcome we aim to predict.

### Models to Follow

```{r}

```


## References

